---
title: "Analysis"
author: "Yeji Sohn"
date: '2023-03-03'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(fig.height = 4, 
    fig.width = 7,
    fig.align   = 'center',
    echo        = FALSE,     
    message     = FALSE, 
    warning     = FALSE,
    digits = 3)
suppressPackageStartupMessages(library(dplyr))
suppressPackageStartupMessages(library(tidyr))
suppressPackageStartupMessages(library(purrr))
suppressPackageStartupMessages(library(plm))
suppressPackageStartupMessages(library(gplots))
suppressPackageStartupMessages(library(glmmLasso))
suppressPackageStartupMessages(library(rempsyc))
suppressPackageStartupMessages(library(ggplot2))
suppressPackageStartupMessages(library(lmtest))
suppressPackageStartupMessages(library(car))
suppressPackageStartupMessages(library(stargazer))
suppressPackageStartupMessages(library(sandwich))
suppressPackageStartupMessages(library(panelAR))
suppressPackageStartupMessages(library(aTSA))
```

# Data Summary
```{r}
file_path = "agg.csv"
agg_df <- read.csv(file_path)
head(agg_df)
```
## Distribution of variables of interest
```{r}
hist(agg_df$REMOTE)
```

```{r}
hist(agg_df$INCOME)
```
Both the dependent and independent variable has strongly positively skewed distribution and normal distribution cannot be assumed.

## Descriptive statistics of variables
```{r}
# Summary of each variables
pre_pandemic_summary <- agg_df %>% 
  filter(YEAR %in% c(2018, 2019))%>%
  summarise_all(mean) %>%
  select(-c(PUMA, YEAR))

pandemic_summary <- agg_df %>% 
  filter(YEAR %in% c(2020, 2021))%>%
  summarise_all(mean) %>%
  select(-c(PUMA, YEAR))

mean_summary <- rbind(pre_pandemic_summary, pandemic_summary) %>% 
  t() %>% 
  as.data.frame() %>%
  rename("Pre-Pandemic_mean" = V1, "Pandemic_mean" = V2)

mean_summary <- cbind(var = rownames(mean_summary), mean_summary)
rownames(mean_summary) <- 1:nrow(mean_summary)

pre_pandemic_summary <- agg_df %>% 
  filter(YEAR %in% c(2018, 2019))%>%
  summarise_all(sd) %>%
  select(-c(PUMA, YEAR))

pandemic_summary <- agg_df %>% 
  filter(YEAR %in% c(2020, 2021))%>%
  summarise_all(sd) %>%
  select(-c(PUMA, YEAR))

sd_summary <- rbind(pre_pandemic_summary, pandemic_summary) %>% 
  t() %>% 
  as.data.frame() %>%
  rename("Pre-Pandemic_sd" = V1, "Pandemic_sd" = V2)

sd_summary <- cbind(var = rownames(sd_summary), sd_summary)
rownames(sd_summary) <- 1:nrow(sd_summary)

t4 <- left_join(mean_summary, sd_summary, by="var")
t4 <- t4[,c(1,2, 4, 3, 5)]

colnames(t4) <- c("Var", "T1.mean", "T1.sd", "T2.mean", "T2.sd")

table4_p <- nice_table(t4, separate.header = TRUE, italics = seq(t4),
                       title = c("Table 4", "Summary"))

table4_p

#flextable::save_as_docx(table4_p, path = "Tables/table4.docx")        
```


```{r}
agg_df$YEAR <- as.factor(agg_df$YEAR)
agg_df$PUMA <- as.factor(agg_df$PUMA)
```

## Heteroskedasticity Test

Breusch–Pagan test is performed to detect heteroskedasticity
```{r}
test <- agg_df %>%
  mutate(EDUC2 = EDUC^2)
m <- lm("log(INCOME) ~ EDUC", test)
bptest(m)

agg_df <- agg_df %>%
  mutate(INCOME_LOG = log(INCOME))

# all variables
x_col <- c("REMOTE", "AGE", "HRSWORK", "EDUC", "EDUC", "FEMALE")
h_test <- function(colname){
  m <- lm(paste("INCOME ~ ", colname, sep=""), agg_df)
  test.out <- bptest(m)
  de <- data.frame(x = c(colname), 
                    BP = c(test.out$statistic), 
                    p.val = c(test.out$p.value), 
                   df=c(test.out$parameter))
  return(de)
}
result <- lapply(x_col, h_test)
bind_rows(result, .id = "column_label")
```
Null Hypothesis: Homoscedasticity is present

Alternative Hypothesis: Heteroscedasticity is present

## Test multicolinearity
```{r, include=FALSE}
# Create formula
race_col <- colnames(agg_df[, grep("RACE", colnames(agg_df))])[-1]
occ_col <- colnames(agg_df[, grep("OCCSOC", colnames(agg_df))])[-1]

r_formula <- paste(race_col, collapse = '+')
o_formula <- paste(occ_col, collapse = '+')
cov_formula <- paste0(r_formula, '+', o_formula)

formula_lin <- paste0("INCOME_LOG ~ REMOTE + AGE + FEMALE + HRSWORK + EDUC", 
                      "+", cov_formula)
```

```{r}
# Multicolinearity test
ols <-lm(formula_lin, data=agg_df)
vif_val <- car::vif(ols)
vif_val[vif_val > 5] 

formula_lin <- gsub("OCCSOC53\\+","",formula_lin)

olst <-lm(formula_lin, data=agg_df)

vif_val <- car::vif(olst)
vif_val[vif_val > 5] 
```

Multicolinearity between occupations are tested. Severe multicolinearity found in OCCSOC11, OCCSOC53, OCCSOC15, and OCCSOC47.

## Test stationarity
```{r}
# Stationarity test on original data
y <- data.frame(split(agg_df$INCOME, agg_df$PUMA))
purtest(y, test = "hadri", exo = "intercept", lags = "AIC")

# Stationarity test on Stochastic (difference from lag value) transformation
diff_df1 <- agg_df %>%
  group_by(PUMA) %>%
  mutate(INCOME_DIFF = INCOME_LOG - dplyr::lag(INCOME_LOG, n=1), REMOTE_DIFF = REMOTE - dplyr::lag(REMOTE, n=1)) %>%
  filter(YEAR != 2018)

y_stochastic <- data.frame(split(diff_df1$INCOME_DIFF, diff_df1$PUMA))
purtest_stochastic <- purtest(y_stochastic, test = "hadri", exo = "intercept", lags = "AIC")

# Stationarity test on Deterministic (difference from model fitting) transformation
diff_df2 <- agg_df %>%
  mutate(REMOTE_LOG = log(REMOTE + 1))

detrend_fun <- function(x){
  # detrend using linear regression
  trend <- lm(x ~ 1 + YEAR, diff_df2)
  dt <- x - predict(trend)
  colname <- deparse(substitute(x))
  diff_df2[[colname]] <- dt
}

puma_year <- agg_df %>% select(PUMA, YEAR)
detrend <- as.data.frame(lapply(diff_df2, detrend_fun)) %>% select(-c(PUMA, YEAR))
diff_df2 <- cbind(puma_year, detrend)


y_deterministic <- data.frame(split(diff_df2$REMOTE, diff_df2$PUMA))
purtest_deterministic <- purtest(y_deterministic, test = "hadri", exo = "intercept", lags = "AIC")

plotmeans(INCOME_LOG ~ YEAR, main="INCOME Trend - Stochastic", data=diff_df1)
plotmeans(REMOTE ~ YEAR, main="Remote-work Trend - Stochastic", data=diff_df1)

plotmeans(INCOME_LOG ~ YEAR, main="INCOME Trend - Deterministic", data=diff_df2)
plotmeans(REMOTE ~ YEAR, main="Remote-work Trend - Deterministic", data=diff_df2)

purtest_stochastic
purtest_deterministic

diff_df <- diff_df1
```
No need to test unit-root since T is very small. Detrended data using Stochastic process will be used.

## Conintegration test
```{r}
coin <- coint.test(agg_df$INCOME, agg_df$REMOTE, d = 0, nlag = NULL, output = TRUE)
coin
```

There is no cointegration for lag = 1 and lag = 2. The detrended data from the previous section will be used for non-fixed effect models. 


# Models

## Pooled OLS

$$
\begin{aligned}
&\text{INCOME}_{i} = \alpha+\beta_{1}*\text{REMOTE}_{i}+\sum_{r}\beta_{2,r}*\text{RACE}_{i}+\sum_{o}\beta_{3,o}\text{OCCSOC}_{i}+\beta_{4}\text{FEMALE}_{i}+\beta_{5}\text{AGE}_{i}+\epsilon_{i}\\
\end{aligned}
$$



```{r, include = FALSE}
ols <-lm(formula_lin, data=diff_df)
ols_robust <- coeftest(ols, vcov.=vcovHAC(ols, type='HC0',cluster="time"))
ols_sum <- summary(ols)
```


```{r}
plot(diff_df$REMOTE, diff_df$INCOME, pch=19, xlab="remote", ylab="average INCOME")
abline(ols,lwd=3, col="red")
```


## Fixed effect regression

Assumption: Linear additive model

### Heterotenity across year
```{r}
plotmeans(INCOME ~ YEAR, main="INCOME Trend across years", data=agg_df)
plotmeans(REMOTE ~ YEAR, main="Remote-work Trend across years", data=agg_df)

```
### Fixed effect with no covariates
$$
\begin{aligned}
&INCOME_{i t}=\lambda_i+\gamma_t+\beta_{1}*REMOTE_{i t}+\epsilon_{i t}\\
\end{aligned}
$$

```{r, include = FALSE}
fixed_nc <- plm("INCOME_LOG ~ REMOTE",
          data = agg_df, effect="twoways", index = c("PUMA", "YEAR"), model = "within", vcov=vcovHC)
fixed_nc_robust <- coeftest(fixed_nc, vcov.=plm::vcovNW(fixed_nc, type='HC0',cluster="time"))
fixed_nc_sum <- summary(fixed_nc)
```


### Fixed effect with covariates

$$
\begin{aligned}
&INCOME_{i t}=\lambda_i+\gamma_t+\beta_{1}*REMOTE_{i t}+
\sum_{r}\beta_{2,r}*\text{RACE}_{i t}+
\sum_{o}\beta_{3,o}\text{OCCSOC}_{i t}+
\beta_{4}\text{FEMALE}_{i t}+\beta_{5}\text{AGE}_{i t}+\epsilon_{i t}\\
\end{aligned}
$$

```{r, include = FALSE}
fixed <- plm(formula_lin,
          data = agg_df, effect="twoways", index = c("PUMA", "YEAR"), model = "within", vcov=vcovHC)
fixed_robust <- coeftest(fixed, vcov.=plm::vcovNW(fixed, type='HC0',cluster="time"))
fixed_sum <- summary(fixed)
```


## Other models

### FGLS

From https://cran.r-project.org/web/packages/plm/plm.pdf

"This framework allows the error covariance structure inside every
group (if effect = "individual", else symmetric) of observations to be fully unrestricted and is
therefore robust against any type of intragroup heteroskedasticity and serial correlation"

```{r}
base<-pdata.frame(diff_df, index = c("PUMA","YEAR"), drop.index = FALSE)
fgls<-pggls(paste0("INCOME_LOG ~ REMOTE + AGE + FEMALE + HRSWORK + EDUC", 
                      "+", cov_formula), model = c("within"), 
           data=base, vcov=vcovHC(type="HC0", cluster="time"))
fgls_robust<- coeftest(fgls)
fgls_sum <- summary(fgls)
```

### AR

```{r}
df_ar<-diff_df %>%
  mutate(YEAR = 2017+as.integer(YEAR)) %>%
  as.data.frame()

AR <- panelAR(formula_lin, 
               data=df_ar, panelVar='PUMA', timeVar='YEAR', autoCorr='psar1',
               panelCorrMethod='pcse', rho.na.rm=TRUE, panel.weight='t-1', bound.rho=TRUE)
AR_robust <- coeftest(AR, vcov.=vcov(AR, type='HC0'))

ar_pred <- predict(AR, df_ar)
rss <- sum((as.numeric(unlist(ar_pred[1])) - df_ar$INCOME_LOG) ^ 2)  ## residual sum of squares
tss <- sum((df_ar$INCOME_LOG - mean(df_ar$INCOME_LOG)) ^ 2)  ## total sum of squares
rsq <- 1 - rss/tss
```

## Estimates

```{r, results = "asis"}
# Stargazer output (with and without RSE)
est <- stargazer(ols, fixed_nc, fixed, fgls_robust, AR_robust, type = "text",
          se = list(NULL, ols_robust[,2], fixed_nc_robust[, 2], 
                    fixed_robust[, 2], 
                     fgls_robust[,2], AR_robust[,2]),
          keep = 'REMOTE',
          column.labels = c("Pooled", "Fixed effect\\\\ & (No covariates)", 
                            "Fixed effect", 
                            "FGLS", "AR"),
          column.sep.width = "-10pt", 
          font.size = "small")
```

```{r}
model_est <- data.frame("Var" = c("REMOTE", "SE", "pval", "Observations", "R2", "Adjusted R2", 
                                  "F Statistic", "F statistic df"),
                        "Pooled" = c(
                          as.numeric(ols$coefficients["REMOTE"]), 
                          ols_robust[,2]["REMOTE"],
                          ols_robust[,4]["REMOTE"],
                          nobs(ols),
                          ols_sum$r.squared,
                          ols_sum$adj.r.squared,
                          ols_sum$fstatistic[1],
                          paste(c("(df = ", ols_sum$fstatistic[2], "; ", ols_sum$fstatistic[3], ")"), collapse = '')
                          ),
                        "Fixed Effect (no covariate)" = c(
                          as.numeric(fixed_nc$coefficients["REMOTE"]),
                          fixed_nc_robust[,2],
                          fixed_nc_robust[,4],
                          nobs(fixed_nc),
                          fixed_nc_sum$r.squared[1],
                          fixed_nc_sum$r.squared[2],
                          fixed_nc_sum$fstatistic$statistic,
                          paste(c("(df = ", fixed_nc_sum$fstatistic$parameter[1], "; ", fixed_nc_sum$fstatistic$parameter[2], ")"), 
                                collapse = '')
                          ),
                        "Fixed Effect" = c(
                          as.numeric(fixed$coefficients["REMOTE"]),
                          fixed_robust[,2]["REMOTE"],
                          fixed_robust[,4]["REMOTE"], 
                          nobs(fixed),
                          fixed_sum$r.squared[1],
                          fixed_sum$r.squared[2],
                          fixed_sum$fstatistic$statistic,
                          paste(c("(df = ", fixed_sum$fstatistic$parameter[1], "; ", fixed_sum$fstatistic$parameter[2], ")"), 
                                collapse = '')
                          ),
                        "FGLS" = c(
                          as.numeric(fgls$coefficients["REMOTE"]),
                          fgls_robust[,2]["REMOTE"],
                          fgls_robust[,4]["REMOTE"], 
                          nobs(fgls),
                          fgls_sum$rsqr,
                          "NA",
                          "NA",
                          paste(c("(df = ", NULL, "; ", NULL, ")"), collapse = '')
                          ),
                        "AR" = c(
                          as.numeric(AR$coefficients["REMOTE"]),
                          AR_robust[,2]["REMOTE"],
                          AR_robust[,4]["REMOTE"],
                          "NA",
                          "NA",
                          "NA",
                          "NA",
                          paste(c("(df = ", NULL, "; ", NULL, ")"), collapse = '')
                          )
                        )

table5_p <- nice_table(model_est,
                       title = c("Table 5", "Estimation"),
                       note = c(""))

# flextable::save_as_docx(table5_p, path = "Tables/table5.docx") 
```

```{r}
# find the average increase in share of remote workers in the sample and to show them what the equivalent in-sample predicted increase in average INCOME would be. 
avg_remote <- agg_df %>%
  group_by(YEAR) %>%
  summarise(mean(REMOTE))

0.073*(0.1967-0.0544)
  
predict(fixed, newdata = avg_remote)
```


# Robustness Checks

```{r}
models <- list(ols, fixed_nc, fixed)
```

#### Testing for fixed effects
```{r}
pFtest(fixed, ols)
```
H0: OLS is better than fixed.

Choose fixed effect over pooled.

#### Hausman test
```{r}
ph_test <- phtest(fixed, random)
```

Choose fixed effect over random effect.


#### Testing for serial correlation
```{r}
durbinWatsonTest(ols)

pdwtest(fixed)
```


```{r}
bg <- sapply(models, function(m) bgtest(m, order=3))
 bgtest(ols, order=3)
```
Since this p-value is less than 0.05, we can reject the null hypothesis and conclude that the residuals in this regression model are perfectly positive autocorrelated.


## Check heteroskedasticity 
```{r}
#load lmtest library
library(lmtest)
bp <- sapply(models, function(m) bptest(m))
bp

bptest(AR)
```

Reject the null hypothesis: heteroscedasticity (Correlation with time) is present in the data. 

```{r}
pcdtest(fixed_nc)
pcdtest(ar)
```

Fail to reject null hypothesis: no cross-sectional dependency in fixed model with covariates. 

```{r}
stationary.test(agg_df$INCOME_LOG, nlag = NULL, output = TRUE)
tseries::adf.test(agg_df$REMOTE)
FinTS::Unitroot(agg_df$REMOTE)
```



# References
https://docs.iza.org/dp13524.pdf
http://web.mit.edu/insong/www/pdf/FEmatch-twoway.pdf
